Synthesizing Quotient Lenses

### 2017-10-18 papers/QuotientLensSynth/main.tex
Checkpoint

In this paper, we address both of these issues.  First, we propose a novel
language of {\em Quotient Regular Expressions}, or QREs, which allows
programmers to specify a broad class of equivalence relations over regular
languages using a compact and convenient notation in which the regular
language and its equivalence classes are captured by a single expression.
We then define {\em QRE lenses}, a class of bijective quotient lenses that
are typed using QREs, and show how to {\em synthesize} QRE lenses from their
QRE types and a set of input-output examples.  We demonstrate the
practicality of this approach by synthesizing QRE lenses between several
real-world data formats from the {\tt data.gov} database.
\end{abstract}

\section{Introduction}
\ifplentyoftime
\bcp{Not sure there is time to fix this, but the intro seems very long and
  technical---it takes a long time to get to saying what our actual
  contributions are!  Wonder if we should move most of it to a second
  ``Example'' section.} 
\fi
{\em Bidirectional programming languages} such as
Boomerang~\cite{boomerang}, Augeas~\cite{augeas}, XSugar~\cite{xsugar},
biXid~\cite{bixid} and X/Inv~\cite{xinv}\bcp{there are many others; we
  should mention all of them} enable a programmer to express a
transformation that maps some input data to some output data, as well as a
transformation that maps the output data back into the input data in a single
expression. An example of a {\em bidirectional transformation} which we shall
revisit severally in this paper is a program which converts citation records
such as
\begin{verbatim}
@Book {conway,
  Author = "J. H. Conway",
  Title = {Regular Algebra and Finite Machines}
}
\end{verbatim}

@@

Bidirectional programs have also been called {\em lenses}. Formally, Foster et
al~\cite{foster} defined a lens $\ell : S \Leftrightarrow V$ that maps between
an ``abstract" {\em source} $S$ and a ``concrete" {\em view} $V$ to be a triple
of functions $(\ell.\get, \; \ell.\lput, \; \ell.\create)$ satisfying the
following equations:\bcp{Why are we showing all the complexity of asymmetric
lenses?  We should just say that we're considering only the bijective case
for now and then just give the details of that.}
\begin{align*}
\ell.get &: S \longrightarrow V & \ell.\lput \; (\ell.\get \; s) \; s &= s\\
\ell.put &: V \longrightarrow S \longrightarrow S & \ell.\get \; (\ell.\lput \;
v \; s) &= v\\
\ell.create &: S \longrightarrow V & \ell.\lput \; (\ell.\create \; v) &= v
\end{align*}

>>>>

In this paper, we address both of these issues.  First, we propose a novel
language of {\em Quotient Regular Expressions}, or QREs, which allows
programmers to specify a broad class of equivalence relations over regular
languages using a compact and convenient notation in which the regular
language and its equivalence classes are captured by a single expression.
We then define {\em QRE lenses}, a class of bijective quotient lenses that
are typed using QREs, and we show how to {\em synthesize} QRE lenses from their
QRE types and a set of input-output examples.  We demonstrate the
practicality of this approach by synthesizing QRE lenses between several
real-world data formats from the {\tt data.gov} database.
\end{abstract}

\section{Introduction}
\ifplentyoftime
\bcp{Not sure there is time to fix this, but the intro seems very long and
  technical---it takes a long time to get to saying what our actual
  contributions are!  Wonder if we should move most of it to a second
  ``Example'' section.} 
\fi
\bcp{This sentence could be clearer!}{\em Bidirectional programming languages} such as
Boomerang~\cite{boomerang}, Augeas~\cite{augeas}, XSugar~\cite{xsugar},
biXid~\cite{bixid} and X/Inv~\cite{xinv}\bcp{there are many others; we
  should mention all of them} enable a programmer to express a
transformation that maps some input data to some output data, as well as a
transformation that maps the output data back into the input data in a single
expression.  An example of a {\em bidirectional transformation} which we shall
revisit severally in this paper is a program which converts citation records
such as
\begin{verbatim}
@Book {conway,
  Author = "J. H. Conway",
  Title = {Regular Algebra and Finite Machines}
}
\end{verbatim}

@@

Bidirectional programs have also been called {\em lenses}. Formally, Foster et
al~\cite{foster} defined a lens $\ell : S \Leftrightarrow V$ that maps between
an ``abstract" {\em source} $S$ and a ``concrete" {\em view} $V$ to be a triple
of functions $(\ell.\get, \; \ell.\lput, \; \ell.\create)$ satisfying the
following equations:\bcp{Why are we showing all the complexity of asymmetric
lenses?  We should just say that we're considering only the bijective case
for now and then just give the details of that.  (Related: The
``abstract/concrete'' terminology doesn't make much sense in the bijective
setting.)} 
\begin{align*}
\ell.get &: S \longrightarrow V & \ell.\lput \; (\ell.\get \; s) \; s &= s\\
\ell.put &: V \longrightarrow S \longrightarrow S & \ell.\get \; (\ell.\lput \;
v \; s) &= v\\
\ell.create &: S \longrightarrow V & \ell.\lput \; (\ell.\create \; v) &= v
\end{align*}
